---
title: "Collaborative Project jfq"
author: "Jorge Quintanilla"
date: "12/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(emmeans)
library(dplyr)
library(ggplot2)
library(esquisse)
```
**Data Cleaning**
```{r}
cdap <- read.table("CollabFile.csv", sep=",",header= TRUE,na.strings = "NA") #read in the file
glimpse(cdap) #check the structure
cdap.cl <- cdap[,1:7] #remove the genotype columns
cdapT<-cdap.cl
glimpse(cdap.cl) #check the removal
```

```{r}
#recode E and W as N and S using map
#south is: elkville, belleville, perry
#north is: hampshire, urbana, neoga
table(cdap.cl$Reg, cdap.cl$Loc) #south/north is not equivalent to east/west
cdap.cl$Reg <- as.character(cdap.cl$Reg)
cdap.cl$Loc <- as.character(cdap.cl$Loc)
cdap.cl[cdap.cl$Loc == "Belleville_2019", "Reg"] <- "south"
cdap.cl[cdap.cl$Loc == "Elkville_2019", "Reg"] <- "south"
cdap.cl[cdap.cl$Loc == "Perry_2019", "Reg"] <- "south"
cdap.cl[cdap.cl$Loc == "Hampshire_2019", "Reg"] <- "north"
cdap.cl[cdap.cl$Loc == "Urbana_2018", "Reg"] <- "north"
cdap.cl[cdap.cl$Loc == "Neoga_2018", "Reg"] <- "north"
table(cdap.cl$Reg, cdap.cl$Loc)
```

```{r}
##Rename the blank treatment
table(cdap.cl$SeedTreatment) #blank variable is supposed to be No treatment
#Outcome was 18 blank variables, and by taking a look at the dataset, there is one "SeedTreatment" in USG_3197
cdap.cl$SeedTreatment <- as.character(cdap.cl$SeedTreatment)
cdap.cl[cdap.cl$SeedTreatment == "", "SeedTreatment"] <- "NTRT"
cdap.cl[cdap.cl$SeedTreatment == "E", ]
cdap.cl[cdap.cl$EntryID == "Becks_Hybrids_Beck_726", ] #this data point is going to cause estimability issues
cleanCollab <- cdap.cl[cdap.cl$SeedTreatment != "E", ]
```

```{r}
#refactor for analysis
cleanCollab$Reg <- as.factor(cleanCollab$Reg)
cleanCollab$Loc <- as.factor(cleanCollab$Loc)
cleanCollab$SeedTreatment <- as.factor(cleanCollab$SeedTreatment)
```

```{r}
#data cleaning for location and region
table(cleanCollab$Loc)
Belleville <- cleanCollab[cleanCollab$Loc == "Belleville_2019", ]
Elkville <- cleanCollab[cleanCollab$Loc == "Elkville_2019", ]
Hampshire <- cleanCollab[cleanCollab$Loc == "Hampshire_2019", ]
Neoga <- cleanCollab[cleanCollab$Loc == "Neoga_2018", ]
Perry <- cleanCollab[cleanCollab$Loc == "Perry_2019", ]
Urbana <- cleanCollab[cleanCollab$Loc == "Urbana_2018", ]
north <- cleanCollab[cleanCollab$Reg == "north", ]
south <- cleanCollab[cleanCollab$Reg == "south", ]
```

**This step provides a long output, not valuable for analysis, just to prove myself It's unbalanced**
```{r}
#model based on data structure
#lm.df <- lm(Estimate ~ Reg/Loc + Company/Variety + SeedTreatment, data = cleanCollab)
#summary(lm.df)
#not really appropriate given the ultimately unbalanced design
```
**Exploratory Analysis*
```{r}
boxplot(cleanCollab$Estimate~cleanCollab$Loc, xlab="location", ylab="Yield", main="Yield by Location")
###Run this line together
hist(cleanCollab$Estimate ,main="Density plot of Yield", prob=T, col="navy")
lines(density((cleanCollab$Estimate)), col="orange", lwd=3)
```


```{r}
ggplot(north) +
 aes(x = Loc, y = Estimate, fill = Company) +
 geom_boxplot() +
 scale_fill_hue() +
 theme_minimal()

ggplot(data=north, aes(x=Estimate, group=Company))+
   geom_density(adjust=1.5, fill= "tomato") +
   facet_wrap(~Company) +
   xlab("Yield estimate") +
   theme(legend.position = "none",
         panel.spacing = unit(0.1, "lines"),
         axis.ticks.x=element_blank())
```


```{r}
ggplot(south) +
 aes(x = Loc, y = Estimate, fill = Company) +
 geom_boxplot() +
 scale_fill_hue() +
 theme_minimal()

ggplot(data=south, aes(x=Estimate, group=Company))+
   geom_density(adjust=1.5, fill= "tomato") +
   facet_wrap(~Company) +
   xlab("Yield estimate") +
   theme(legend.position = "none",
         panel.spacing = unit(0.1, "lines"),
         axis.ticks.x=element_blank())
```
**MANY OF THE LS MEANS COMPARISON PLOTS WERE UNREADABLE, HENCE, THEY WERE NOT INCLUDED IN THE FINAL REPORT, INSTEAD A TABLE WITH SUMMARIZED RESULTS WAS REPORTED**
```{r}
###Objective 1. Evaluate Location effect
mylm1 <- lm(Estimate ~ Loc, data = cleanCollab)
anova(mylm1)
summary(mylm1) #r2 Loc = 0.79
plot(mylm1)
AvLoc <- emmeans(mylm1, specs = "Loc", adjust = "tukey")
plot(AvLoc, comparisons = TRUE, horizontal = FALSE)
pwpp(AvLoc, comparisons = TRUE)
```

```{r}
###Objective 2: Evaluate Company effect
mylm2 <- lm(Estimate ~ Company, data = cleanCollab)
summary(mylm2) #some companies have significant relationshps with yield
anova(mylm2) #overall the effect of company is significant

OvPerf<- cleanCollab%>%
  group_by(Company, Variety)%>%
  summarize(mean=mean(Estimate))
OvPerf 
#The outcome of this is DeRaedt_Seed,DeRaedt_24,106.62373 as the best
#The outcome of this is Go_Wheat,EXP18-1,79.82095 as the worst

#Across regions
CompMean<-plot(emmeans(mylm2, specs = "Company", adjust = "tukey"), comparisons = TRUE)
CompMean #The best is DeRaedt_Seed since the variability is low, 
         #the worst seems to be GoWheat as well, since the mean and extreme values indicates underperformance compared to other companies, that may have higher upper limit and mean 
CompMean2<-pwpp(emmeans(mylm2, specs = "Company", adjust = "tukey"), comparisons = TRUE)
CompMean2 #This is for the significance of each company
table(cdap.cl$Reg, cdap.cl$Company) ### Some companies belong in different regions
TukeyHSD(aov(Estimate ~ Company, data = cleanCollab))
```

```{r}
#lm of each location
Bellm <- lm(Estimate ~ Company, data = Belleville)
Elklm <- lm(Estimate ~ Company, data = Elkville)
Hamplm <- lm(Estimate ~ Company, data = Hampshire)
Neolm <- lm(Estimate ~ Company, data = Neoga)
Pelm <- lm(Estimate ~ Company, data = Perry)
Urblm <- lm(Estimate ~ Company, data = Urbana)

##Run the ANOVA for the lm for each location
anova(Bellm) #no sig
anova(Elklm) #sig
anova(Hamplm) #sig
anova(Neolm) #no sig
anova(Pelm) #no sig
anova(Urblm) #sig
summary(Bellm)
```

```{r}
OvPerfLoc2<- cleanCollab%>%
  group_by(Loc, Company)%>%
  summarize(mean=mean(Estimate))%>%
  arrange()
summarise(OvPerfLoc2, max(mean))
summarise(OvPerfLoc2, min(mean))
#Worst/Best Company in each location
#Belleville= ProHarvest/Pioneer
#Elkville= ProHarvest/KWS_Cereals
#Hampshire= Kratz_Farms/Pioneer
#Neoga= Miller_Bros_Farm_and_Fert/Croplan
#Perry= ProHarvest/Pioneer
#Urbana=Kratz_Farms/Pioneer
```

```{r}
##Calculate the  lsMeans
MBelleville <- emmeans(Bellm, specs = "Company", adjust = "tukey")
MElkville <- emmeans(Elklm, specs = "Company", adjust = "tukey")
MHampshire <- emmeans(Hamplm, specs = "Company", adjust = "tukey")
MNeoga <- emmeans(Neolm, specs = "Company", adjust = "tukey")
MPerry <- emmeans(Pelm, specs = "Company", adjust = "tukey")
MUrbana <- emmeans(Urblm, specs = "Company", adjust = "tukey")


##Plot the lsMeans
pwpp(MBelleville, comparisons = TRUE, ylab = "Company of Belleville")
pwpp(MElkville, comparisons = TRUE, ylab = "Company of Elkville")
pwpp(MHampshire, comparisons = TRUE, ylab = "Company of Hampshire")
pwpp(MNeoga, comparisons = TRUE, ylab = "Company of Neoga")
pwpp(MPerry, comparisons = TRUE, ylab = "Company of Perry")
pwpp(MUrbana, comparisons = TRUE, ylab = "Company of Urbana")
```

```{r}
###Objective 3: Evaluate the variety 
#variety across region
mylm3 <- lm(Estimate ~ Variety, data = cleanCollab)
summary(mylm3) #level of region has sig effect on yield
anova(mylm3) #variety does not have sig effect on yield
plot(mylm3)
plot(emmeans(mylm3, specs = "Variety", adjust = "tukey"), comparisons = TRUE)
pwpp(emmeans(mylm3, specs = "Variety", adjust = "tukey"), comparisons = TRUE)
```


```{r}
VarPerf<-cleanCollab%>%
  group_by(Reg, EntryID)%>%
  summarize(mean=mean(Estimate))%>%
  arrange(mean)

VarPerf2<-cleanCollab%>%
  group_by(EntryID)%>% #In EntryID There is both the company and the variety 
  summarise(mean=mean(Estimate))%>%
  arrange(mean)

VarPerf3<-cleanCollab%>%
  group_by(Variety)%>%
  summarise(mean=mean(Estimate))%>%
  arrange(mean)

summarise(VarPerf3, max(mean))
summarise(VarPerf2, max(mean))
summarise(VarPerf2, min(mean))
#Worst/Best  varieties across regions
#Go_Wheat and DeRaedt_Seed
#####Check
#North= Hoffman_Seed var. H7W17/KWS_Cereals var. KWS19X03
#South= Proharvest Var. 317/Croplan var. CP9606
```
These linear models and ANOVA were performed, but the main result was a perfect fit, which is unreliable. However, variety was significant at each location accordingly to the analysis of variance.

```{r}
#variety by location
#vlm <- lm(Estimate ~ Variety + Loc, data = cleanCollab)
#vBellm <- lm(Estimate ~ Variety, data = Belleville)
#vElklm <- lm(Estimate ~ Variety, data = Elkville)
#vHamplm <- lm(Estimate ~ Variety, data = Hampshire)
#vNeolm <- lm(Estimate ~ Variety, data = Neoga)
#vPelm <- lm(Estimate ~ Variety, data = Perry)
#vUrblm <- lm(Estimate ~ Variety, data = Urbana)

#anova(vlm)
#anova(vBellm) #sig
#anova(vElklm) #sig
#anova(vHamplm) #sig
#anova(vNeolm) #sig
#anova(vPelm) #sig
#anova(vUrblm) #sig
```

```{r}
OvPerfBel<-Belleville %>%
  group_by(Loc, EntryID)%>%
  summarize(mean=mean(Estimate))%>%
  arrange(mean)
OvPerfBel

OvPerfElk<- Elkville%>%
  group_by(Loc, EntryID)%>%
  summarize(mean=mean(Estimate))%>%
  arrange(mean)
OvPerfElk
OvPerfHamp<- Hampshire%>%
  group_by(Loc, EntryID)%>%
  summarize(mean=mean(Estimate))%>%
  arrange(mean)
OvPerfHamp
OvPerfNeo<- Neoga%>%
  group_by(Loc, EntryID)%>%
  summarize(mean=mean(Estimate))%>%
  arrange(mean)
OvPerfNeo
OvPerfPe<- Perry%>%
  group_by(Loc, EntryID)%>%
  summarize(mean=mean(Estimate))%>%
  arrange(mean)
OvPerfPe
OvPerfUrb<- Urbana%>%
  group_by(Loc, EntryID)%>%
  summarize(mean=mean(Estimate))%>%
  arrange(mean)
OvPerfUrb

#Worst/Best Variety in each location
#Belleville= ProHarvest_317/KWS_Cereals_KWS19X07
#Elkville= ProHarvest_286/KWS_Cereals_KWS19X09
#Hampshire= Kratz_Farms_KF_15334/KWS_Cereals_KWS19X03
#Neoga= GROWMARK_FS_604/Hoffman_Seed_H7W18
#Perry= Kratz_Farms_KF_15334/AgriMAXX_495
#Urbana=LEWIS_851/GROWMARK_WX18C

```

```{r}
###Objective 4: Evaluate Regional Effects
mylm4 <- lm(Estimate ~ Reg, data = cleanCollab)
summary(mylm4) #Region alone explains 16% of variance
anova(mylm4)
mylm4.1 <- lm(Estimate ~ Reg + Company, data = cleanCollab)
anova(mylm4.1)
summary(mylm4.1) #Together with company explain 19% of variance
```

```{r}
Northlm <- lm(Estimate ~ Company, data = north)
Southlm <- lm(Estimate ~ Company, data = south)

MNorth <- emmeans(Northlm, specs = "Company", adjust = "tukey")
MSouth <- emmeans(Southlm, specs = "Company", adjust = "tukey")
pwpp(MNorth, comparisons = TRUE) ##Six or seven of them seem sig for North Reg
pwpp(MSouth, comparisons = TRUE) ##Seems not sig for SOUTH
##Best Company within Region= N:Moiner_Seed/S:DeRaedt_Seed
##Worst Company within Region= N= Miller_Bros_Farm_and_Fert/S: Proharvest

```


```{r}
#linear model for north/south company's variety
NorthVarlm <- lm(Estimate ~ Variety, data = north)
SouthVarlm <- lm(Estimate ~ Variety, data = south)
MNorthVar <- emmeans(NorthVarlm, specs = "Variety", adjust = "tukey")
MSouthVar <- emmeans(SouthVarlm, specs = "Variety", adjust = "tukey")
pwpp(MNorthVar, comparisons = TRUE) #No sig for north/south
pwpp(MSouthVar, comparisons = TRUE)
#KWS19X03 Best in North/H7W17 Worst in North 
#CP9606 Best in South/ 317 Worst in South
```

```{r}
###Objective 5: Evaluate Seed Treatment 
#lm for ST = significant at 0.03
mylm5 <- lm(Estimate ~ SeedTreatment, data = cleanCollab)
anova(mylm5) #significant
pwpp(emmeans(mylm5, specs = "SeedTreatment", adjust = "tukey"), comparisons = TRUE)
TukeyHSD(aov(Estimate ~ SeedTreatment, data = cleanCollab))
summary(mylm5) #Significant for no treatment, however, Treatment C has the higher positive effect, but it is not significant
```

```{r}
#Model Selection - Using stepwise selection
test_int <- lm(Estimate~1, data=cleanCollab)
test_full <- lm(Estimate~Reg+Loc+Company+Variety+SeedTreatment, data=cleanCollab)
SWmodel<-step(test_int,scope = list(upper = formula(test_full), lower = formula(test_int)), direction = 'both')
summary(SWmodel)$terms
### Best model= Estimate~Loc+Variety, data=cleanCollab)
```

```{r}
###Objective 6: Predictions
mylm6 <- lm(Estimate ~ Loc + Variety, data = cleanCollab)
summary(mylm6) #Hampshire i the max
summary(mylm6)$coef
max(coef(mylm6)[grep("Variety*", names(coef(mylm6)))]) #VarietyH7W18
##Hampshire and variety H7W8 are the best for maximum yield overall
mylm6.1 <- lm(Estimate ~ Loc + Variety+ SeedTreatment, data = cleanCollab)
summary(mylm6.1) #Not justified for treat the seed.
```

```{r}
mylm7<-lm(Estimate~Loc, data=north)
NorthLoc<-emmeans(mylm7, specs = "Loc", adjust = "tukey") #Hampshire is best location in North
plot(NorthLoc, comparisons = TRUE, horizontal = FALSE)
##We know the best variety of the North = KWS_Cereals var. KWS19X03
```

```{r}
mylm8<-lm(Estimate~Loc, data=south)
SouthLoc<-emmeans(mylm8, specs = "Loc", adjust = "tukey") #Elkville is best location in South
plot(SouthLoc, comparisons = TRUE, horizontal = FALSE)
##We know the best variety of the South = Croplan var. CP9606
```

```{r}
#Predictions for North
Northmylm6.2 <- lm(Estimate ~ Loc + Variety, data = north)
BNorth<-data.frame(Loc="Hampshire_2019", Variety="KWS19X03")
predict(Northmylm6.2, BNorth)
```

```{r}
#Predictions for South
Southmylm6.2 <- lm(Estimate ~ Loc + Variety, data = south)
BSouth<-data.frame(Loc="Elkville_2019", Variety="CP9606")
predict(Southmylm6.2, BSouth)
```





